Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis

### 2021-05-12 writing /thesis_draft/index.Rmd
final copy submitted to the library

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g. El Niño Southern Oscillation).  As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems [@khelifa_usefulness_2019; @stewart_mesocosm_2013]. How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].

## Why this study on variability matters 
There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g. inflationary effects in sink-source systems [@holt_impacts_2003; @vasseur_phase-locking_2009], such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, e.g. individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions. We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response. To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

>>>>

## Relevant significance of variability today 
Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g. El Niño Southern Oscillation).  As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems [@khelifa_usefulness_2019; @stewart_mesocosm_2013]. How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].

## Why this study on variability matters 
There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g. inflationary effects in sink-source systems [@holt_impacts_2003; @vasseur_phase-locking_2009], such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, i.e. individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions. We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response. To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.
